\bf Comparing the Power of Plot Designs to Reveal Correlation

### 2015-12-17 paper/paper.Rnw
Rewrite model predict

\begin{table}[htp]
\caption{Weighted Mixed Effects Logit Model for the predicted power of plot design. Sample size for scatter plot model: 900, overlaid lines model: 949, total: 1849. Default plot design is an overlaid lines type. The model is fitted using the glimmix procedure in SAS. Subjects with less than ten lineup evaluations were removed from the sample. %Sign = 1 if correlation > 0, 0 otherwise. Each plot design is estimated separately. Results must be transformed via the logit link function to generate a power prediction. Subjects with less than ten lineup evaluations were removed from the sample..}
\label{tab1}
\centering
<<powerfit, dependson='setup', results='asis', echo=FALSE, message=FALSE, error=FALSE, warning=FALSE>>=
full$subj_id <- factor(full$ip_address, labels=1:207)

# Need to keep only evaluations done by subject who completed at least 10
x <- data.frame(table(full$subj_id))
keep <- x$Var1[x$Freq>9]
full.sub <- filter(full, subj_id %in% keep)
full.sub$r <- round(full.sub$correlation, 1)
full.sub$rsgn <- ifelse(full.sub$r >= 0, 1, 0)
full.sub$rabs <- abs(full.sub$r)
#full.sub$rabs <- factor(full.sub$rabs)
full.sub$plot <- ifelse(full.sub$test_param == 0, "S", "L")
full.sub <- select(full.sub, subj_id, correct, r, rsgn, rabs, smoothed, n, plot, log_time, response_0, actual_location)#[,c(1,7, 15,16,17,20:26)]
full.sub$attempts <- sapply(strsplit(as.character(full.sub$response_0), ","), length)
full.sub$weights = 1/(full.sub$attempts) #weights are actually number of attempts in the glm
full.sub <- full.sub %>% transform(response_0= strsplit(as.character(response_0), ",")) %>% unnest(response_0)
full.sub$correct <- ifelse(full.sub$response_0 == full.sub$actual_location, 1, 0) #fixing correct for multiple guesses
#full.sub$rpos <- ifelse(full.sub$r > 0, full.sub$r, 0)
#full.sub$rneg <- ifelse(full.sub$r < 0, full.sub$r, 0)
full.sub$alpha <- log(.05/0.95)
#full.sub$correct.prop <- full.sub$correct/full.sub$attempts #correct.prop - proportion of correct selections, weights - number of attempts
#weirw.csv(full.sub, "choices.csv") #exporting data to be estimated in SAS.
#m8l <- glmer(correct.prop~offset(alpha) + rabs*rsgn + smoothed + n + (1|subj_id), data=filter(full.sub, plot == "L"), family="binomial", weights=attempts)
#m8s <- glmer(correct.prop~offset(alpha) + rabs*rsgn + smoothed + n + (1|subj_id), data=filter(full.sub, plot == "S"), family="binomial", weights=attempts)
#pander(summary(m8l), justify=c('left', 'right', 'right', 'right', 'right', caption="Model fit for overlaid lines"))
#pander(summary(m8s)$coefficients, justify=c('left', 'right', 'right', 'right', 'right'), caption="Model fit for scatter plots")
#library(xtable)
#line.fit <- summary(m8l)$coefficients
#scatter.fit <- summary(m8s)$coefficients
#colnames(scatter.fit)[4] <- "p-value"
#rownames(scatter.fit) <- c("Intercept", "abs(r)", "sign(r)", "smoothed", "t", "Interaction(r+, r-)")
#colnames(line.fit)[4] <- "p-value"
#rownames(line.fit) <- c("Intercept", "abs(r)", "sign(r)", "smooth", "sample size", "Interaction(r+, r-)")
#kable(scatter.fit, format="latex", digits=2, caption="Scatter plot")
#kable(line.fit, format="latex", digits=2, caption="Overlaid Lines")
@
\end{table}

@@

\begin{figure}[htbp]
 \centering
<<powermodel, dependson='powerfit', fig.height=6, fig.width=6, out.width='0.7\\textwidth', echo=FALSE>>=
r <- seq(-0.9,0.9,0.02)
n <- c(12, 24, 48, 96)
predl <- data.frame(expand.grid(r=r, n=n, alpha=log(.05/0.95),
                                      smoothed=c(0,1), 
                                      subj_id=subset(full.sub, plot=="L")$subj_id)) #ranef(m4)[[1]][,1]))
predl$rsgn <- ifelse(predl$r > 0, 1, 0)
predl$rabs <- abs(predl$r)
predl$fitre <- predict(m8l, predl, type="response", re.form=~(1|subj_id),
                             allow.new.levels=TRUE)
predl$fitfx <- predict(m8l, predl, type="response", re.form=NA,
                             allow.new.levels=TRUE)
predl$plot <- ifelse(predl$smoothed == 0, "Overlaid Lines", "Overlaid Lines (Smoothed)")
predl <- filter(predl, smoothed == 0 | smoothed == 1 & n==48 | smoothed ==1 & n==96)

preds <- data.frame(expand.grid(r=r, n=n, alpha=log(.05/0.95),
                                      smoothed=c(0,1), 
                                      subj_id=subset(full.sub, plot=="S")$subj_id)) #ranef(m4)[[1]][,1]))
preds$rsgn <- ifelse(preds$r > 0, 1, 0)
preds$rabs <- abs(preds$r)
preds$fitre <- predict(m8s, preds, type="response", re.form=~(1|subj_id),
                             allow.new.levels=TRUE)
preds$fitfx <- predict(m8s, preds, type="response", re.form=NA,
                             allow.new.levels=TRUE)
preds$plot <- ifelse(preds$smoothed == 0, "Scatterplot", "Scatterplot (Smoothed)")
preds <- filter(preds, smoothed == 0 | smoothed == 1 & n==48 | smoothed ==1 & n==96)
full.pred <- rbind(preds, predl)

#full.pred <- mutate(full.pred2, plot2 = ifelse(substr(plot, 1, 1) =="S", "Scatterplot", "Overlaid Lines"))

qplot(r, fitre, data=full.pred, color=plot, 
      group=interaction(subj_id, plot), geom="line", alpha=I(0.1)) + 
  scale_color_manual("Plot Design", values=c("#33AA33", "#9977EE", "#FF9933", "#00DDAA")) + 
  xlab("Correlation") + ylab("Estimated Power") + labs(title="Predicted Power") +
  geom_line(aes(y=fitfx), size=2, alpha=1) + theme_bw() + facet_wrap(~n, ncol = 2) + theme(aspect.ratio=1.2)
@
  \label{powermodel}
  \caption{Predicted power fit from the mixed effects logit power model. The weakness of the overlaid line graph in negative correlation and  the scatter plot's gain from sample size are highlighted.}
\end{figure}

\begin{figure}
  \centering
<<residuals, dependson='powerfit', fig.height=4, fig.width=6, out.width='0.7\\textwidth', echo=FALSE, warning=FALSE>>=
resids <- resid(m8s)
residl <- resid(m8l)
resids[842:843] <- NA
resid <- data.frame(line = residl, scatter = resids)
p1b <- ggplot(data=resid[1:841,]) + geom_histogram(aes(y=..density.., x=scatter)) + labs(title="Scatter plot", x="residual", y=element_blank())
p2b <- ggplot(data=resid) + geom_histogram(aes(y=..density.., x=line))+ labs(title="Overlaid Lines", x="residual", y=element_blank())
grid.arrange(p1b, p2b, ncol=2, top = "Residual density plots")
@
  \label{residuals}
  \caption{Histograms of residuals for both the scatter plot (left) and the overlaid lines (right).}
\end{figure}

@@

\begin{table}
\label{tab2}
\centering
\caption{Mixed Linear Model fit for time taken to make a selection}
<<timefit, dependson='powerfit', echo=FALSE, results='hide', message=FALSE, error=FALSE, warning=FALSE>>=
timelmer <-  lmer(log_time ~ smoothed + r + n + rsgn*r + plot + plot*smoothed 
                  + plot*r + plot*n + plot*rsgn*r + plot*correct + (1|subj_id), data=full.sub)
time.fit <- summary(timelmer)$coefficients
pv <- function(x){
  p <- 2*(1-pt(x, 1835))
}
p.value <- sapply(time.fit[,3], pv)
time.fit <- cbind(time.fit, p.value)
time.fit <- time.fit[c(1, 2, 4, 3, 5, 8, 7, 6, 9, 11, 10, 12, 14, 13),] #separate into main effects then interaction effects in same order
rownames(time.fit) <- c("Intercept", "Smoothed", "t", "r", "sign(r)", "r*sign(r)", "correct", "Plot", "Plot*Smoothed", "Plot*t", "Plot*r", "Plot*sign(r)", "Plot*r*sign(r)", "Plot*correct")
kable(time.fit[1:7,], format="latex", caption="Main Effects (Overlaid Lines)")
kable(time.fit[8:14,], format="latex", caption="Interaction Effects (Scatter plot)")
@
\end{table}

>>>>

\begin{table}[htp]
\caption{Weighted Mixed Effects Logit Model for the predicted power of plot design. Sample size for scatter plot evaluations: 900, overlaid lines: 949, total: 1849. Default plot design is an overlaid lines type. The model is fitted using the glimmix procedure in SAS. Subjects with less than ten lineup evaluations were removed from the sample. %Sign = 1 if correlation > 0, 0 otherwise. Each plot design is estimated separately. Results must be transformed via the logit link function to generate a power prediction. Subjects with less than ten lineup evaluations were removed from the sample..}
\label{tab1}
\centering
<<powerfit, dependson='setup', results='asis', echo=FALSE, message=FALSE, error=FALSE, warning=FALSE>>=
full$subj_id <- factor(full$ip_address, labels=1:207)

# Need to keep only evaluations done by subject who completed at least 10
x <- data.frame(table(full$subj_id))
keep <- x$Var1[x$Freq>9]
full.sub <- filter(full, subj_id %in% keep)
full.sub$r <- round(full.sub$correlation, 1)
full.sub$rsgn <- ifelse(full.sub$r >= 0, 1, 0)
full.sub$rabs <- abs(full.sub$r)
full.sub$plot <- ifelse(full.sub$test_param == 0, "S", "L")
full.sub <- select(full.sub, subj_id, correct, r, rsgn, rabs, smoothed, n, plot, log_time, response_0, actual_location)#[,c(1,7, 15,16,17,20:26)]
full.sub$attempts <- sapply(strsplit(as.character(full.sub$response_0), ","), length)
full.sub$weights = 1/(full.sub$attempts) #weights are actually number of attempts in the glm
full.sub <- full.sub %>% transform(response_0= strsplit(as.character(response_0), ",")) %>% unnest(response_0)
full.sub$correct <- ifelse(full.sub$response_0 == full.sub$actual_location, 1, 0) #fixing correct for multiple guesses
full.sub$alpha <- log(.05/0.95)
full.sub$subj_id <- as.numeric(full.sub$subj_id)
subsetted <- select(full.sub, rabs, n, alpha, rsgn, smoothed, plot, subj_id, weights, correct)
#write.csv(subsetted, "sas/choices.csv")

powercoef <- read.csv("sas/PowerEstimates.csv")
powercoef <- powercoef[,-1]
rownames(powercoef) <- c("Intercept", "abs(r)", "sign(r)", "abs(r)*sign(r)", "smoothed", "t", "plot", "plot*abs(r)", "plot*sign(r)", "plot*abs(r)*sign(r)", "plot*smoothed", "plot*t")
kable(powercoef[1:6,], format="latex", digits=2, caption="Main Effects")
kable(powercoef[7:12,], format="latex", digits=2, caption="Interaction Effects")
@
\end{table}

@@

\begin{figure}[htbp]
 \centering
<<powermodel, dependson='powerfit', fig.height=6, fig.width=6, out.width='0.7\\textwidth', echo=FALSE>>=
rabs <- seq(0,0.9,0.02)
t <- c(12, 24, 48, 96)
#pred <- data.frame(expand.grid(rabs=rabs, t=t, alpha=log(.05/0.95), rsgn=c(0,1),
#                                      smoothed=c(0,1), plot=c(0,1),
#                                      subj_id=full.sub$subj_id))
#write.csv(pred, "sas/predgrid.csv") # these lines are done, but its a large file so commented out for knitr
#Code exported to sas, see sas/model\ fit.sas
prediction <- read.csv("sas/predfit.csv") #Large file, very slow
rownames(prediction) <- c("rabs", "t", "alpha", "rsgn", "smoothed", "plot", "subj_id", "weights", "detected", "RandomEf", "FixedEf", "resid")
resid <- subset(prediction, detected==0 | detected==1)
prediction <- mutate(prediction, r = ifelse(rsgn==1, rabs, -rabs), plot = ifelse(plot==1, "Overlaid Lines", "Scatterplot"), plot = ifelse(smoothed==1, paste(plot, "(Smoothed)", sep=" "), plot))
prediction <- select(prediction, rabs, t, rsgn, smoothed, plot, subj_id, RandomEf, FixedEf)
prediction <- mutate(prediction, RandomEf = exp(RandomEf)/(1+exp(RandomEf)), FixedEf = exp(FixedEf)/(1+exp(FixedEf)))

qplot(r, RandomEf, data=prediction, color=plot, 
      group=interaction(subj_id, plot), geom="line", alpha=I(0.1)) + 
  scale_color_manual("Plot Design", values=c("#33AA33", "#9977EE", "#FF9933", "#00DDAA")) + 
  xlab("Correlation") + ylab("Estimated Power") +
  geom_line(aes(y=FixedEf), size=2, alpha=1) + theme_bw() + facet_wrap(~t, ncol = 2) + theme(aspect.ratio=1.2)
@
  \label{powermodel}
  \caption{Predicted power fit from the mixed effects logit power model. The weakness of the overlaid line graph in negative correlation and  the scatter plot's gain from sample size are highlighted.}
\end{figure}

\begin{figure}
  \centering
<<residuals, dependson='powerfit', fig.height=4, fig.width=6, out.width='0.7\\textwidth', echo=FALSE, warning=FALSE>>=
resids <- subset(resid, plot==1)
residl <- subset(resid, plot==0)
#resids[842:843] <- NA  #fix residual stuff later
#resid <- data.frame(line = residl, scatter = resids)
#p1b <- ggplot(data=resid[1:841,]) + geom_histogram(aes(y=..density.., x=scatter)) + labs(title="Scatter plot", x="residual", y=element_blank())
#p2b <- ggplot(data=resid) + geom_histogram(aes(y=..density.., x=line))+ labs(title="Overlaid Lines", x="residual", y=element_blank())
#grid.arrange(p1b, p2b, ncol=2, top = "Residual density plots")
@
  \label{residuals}
  \caption{Histograms of residuals for both the scatter plot (left) and the overlaid lines (right).}
\end{figure}

@@

\begin{table}
\label{tab2}
\centering
\caption{Mixed Linear Model fit for time taken to make a selection, default plot style is overlaid lines.}
<<timefit, dependson='powerfit', echo=FALSE, results='hide', message=FALSE, error=FALSE, warning=FALSE>>=
timelmer <-  lmer(log_time ~ smoothed + r + n + rsgn*r + plot + plot*smoothed 
                  + plot*r + plot*n + plot*rsgn*r + plot*correct + (1|subj_id), data=full.sub)
time.fit <- summary(timelmer)$coefficients
pv <- function(x){
  p <- 2*(1-pt(x, 1835))
}
p.value <- sapply(time.fit[,3], pv)
time.fit <- cbind(time.fit, p.value)
time.fit <- time.fit[c(1, 2, 4, 3, 5, 8, 7, 6, 9, 11, 10, 12, 14, 13),] #separate into main effects then interaction effects in same order
rownames(time.fit) <- c("Intercept", "Smoothed", "t", "r", "sign(r)", "r*sign(r)", "correct", "Plot", "Plot*Smoothed", "Plot*t", "Plot*r", "Plot*sign(r)", "Plot*r*sign(r)", "Plot*correct")
kable(time.fit[1:7,], format="latex", caption="Main Effects", digits=2)
kable(time.fit[8:14,], format="latex", caption="Interaction Effects", digits=2)
@
\end{table}
